---
title: "Attendance Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, warnings = FALSE)
```

## Attendance study

-- collected Cleveland Indians (baseball) attendance for 81 home games in 2016 season

-- response is fraction of capacity

-- input:  game is Weekend/Weekday

-- input:  period of season (first, second, third)

```{r}
library(readr)
d <- read_csv("data/tribe2016.csv")
```

## Beta regression model 

-- Suitable for modeling response data which are rates or proportions

-- $y_i \sim Beta$ with shape parameters $a$ and $b$

-- $a = \mu \phi, \, \, b = (1 - \mu) \phi$

($\mu$ is the mean, $\phi$ is a precision parameter)

-- $logit \mu = \alpha + x \beta$

(logistic model on the means)

## Traditional beta regression

- using function `betareg` in `betareg` package

```{r}
library(betareg)
fit <- betareg(fraction ~ Weekend + Period, data=d,
               link="logit")
```

## Output from betareg

```{r}
summary(fit)
```

## Using STAN in rstanarm package

- Function `stan_betareg` implements MCMC sampling of a Bayesian beta regression model

- Same model syntax as `betareg`

- Can specify a variety of priors (we'll use default one here)

```{r}
library(rstanarm)
fit2 <- stan_betareg(fraction ~ Weekend + Period, data=d,
                     link="logit")
```

## What priors are used?

```{r}
prior_summary(fit2)
```


## MCMC diagnostics -- trace plots

```{r}
library(bayesplot)
mcmc_trace(as.matrix(fit2))
```

## Autocorrelation plots

```{r}
mcmc_acf(as.matrix(fit2))
```

## Density plots for all parameters

```{r}
mcmc_dens(as.matrix(fit2))
```

## Posterior interval estimates

```{r}
posterior_interval(fit2)
```

## Matrix of simulated draws from posterior

```{r}
posterior_sims <- as.matrix(fit2)
head(posterior_sims)
```

## Interested in expected attendance, weekdays, each period

```{r}
library(arm)
d1 <- data.frame(Label="Weekday Period 1", 
      Fraction=invlogit(posterior_sims[, "(Intercept)"]))
d2 <- data.frame(Label="Weekday Period 2", 
      Fraction=invlogit(posterior_sims[, "(Intercept)"] +
                      posterior_sims[, "PeriodSecond"]))
d3 <- data.frame(Label="Weekday Period 3", 
      Fraction=invlogit(posterior_sims[, "(Intercept)"] +
                      posterior_sims[, "PeriodThird"]))
```

## Posteriors of expected attendance, weekday, each period

```{r, fig.height=4}
library(ggplot2)
ggplot(rbind(d1, d2, d3), aes(Fraction)) +
  geom_density() + facet_wrap(~ Label, ncol=1)
```


## Nice graphical interface 

-- Launches graphical interface for diagnostics/summaries

```
launch_shinystan(fit2)
```

## Commands for posterior predictive checking

-- shows density plot of response and some replicated posterior predictive data

```{r, fig.height=3}
pp_check(fit2)
```

## Obtain replicated simulations from posterior predictive distribution

```{r}
ynew <- posterior_predict(fit2)
```

-- need `ynew` to implement the following posterior predictive checks

-- compare $T(y_{rep})$ with $T_{obs}$ using specific checking function $T$

## Posterior predictive check

- using "median" as the test statistic

```{r, fig.height=3}
ppc_stat(d$fraction, ynew, stat="median")
```

## Posterior predictive check

- using "sd" as the test statistic

```{r, fig.height=3}
ppc_stat(d$fraction, ynew, stat="sd")
```

## Posterior predictive check

- using "min" as the test statistic

```{r, fig.height=3}
ppc_stat(d$fraction, ynew, stat="min")
```

## Exercises

1.  Is the beta regression a reasonable model for this attendance data?

2.  Assuming the answer to question 1 is "no", what models might you try next?

3.  What is the advantage of a Bayesian fit compared to a traditional (ML) fit in this setting?

4.  What alternative priors might you try?

5.  (Logistic Modeling)

Suppose you are interested in looking at the relationship between insecticide dose and kill (1 or 0) for 10 insects.  Here is R code setting up the data and implementing a traditional logistic fit.

```
dose <- 1:10
kill <- c(rep(0, 6), rep(1, 4))
df <- data.frame(dose, kill)
fit <- glm(kill ~ dose, 
    family=binomial, data=df)
```

-- This code will produce a warning?  What is going on?  (Look at the parameter estimates and standard errors.)

-- Run this model using the `stan_glm` function.  Compare the parameter estimates for the two fits.  Why are they so different?
