---
title: "BBS Trend Summaries for Grassland Birds"
output:
  output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

## BBS Case Study from Link and Barker's book **Bayesian Inference: With Ecological Applications**

- Data from the North American Breeding Bird Survey (BBS)
- Goal of the yearly survey is to monitor bird population at large geographic scales
- Data collected for over 400  species (focus on 28 grassland species)
- Interested in measuring trend from 1966 to 1999
- Focus on summaries of temporal trend parameters $\beta_i$ for species $1, ..., K$.
- For each of 28 species, have trend estimate $\hat \beta_i$, the corresponding standard error $SE_i$, and the number of sites $n_i$.

## Read in the data

```{r}
library(tidyverse)
(d <- read_csv("data/BBS_survey.csv"))
```

## Derived Parameters of Interest

- How many of these species have negative trends? 

- That is, we are interested in the derived parameter
$$
D(\beta_1, ..., \beta_{28}) = \sum_{i=1}^{28} I(\beta_i < 0)
$$

## Derived Parameters of Interest

- What is the actual rank of the 27th species?

$$
R_{27} = \sum_{i=1}^{28} I(\beta_{27} \ge \beta_i)
$$

## The Model

Sampling Model:

1.  Assume the estimates $\hat \beta_i$ are independent

2.  ${\hat \beta}_i \sim N(\beta_i, 1 / \tau({\hat \beta}_i))$

3.  $SE_i \sim Gamma(n/2, (n/2) \tau({\hat \beta}_i))$

## The Prior (Multilevel)

1. Assume $\beta_1, ..., \beta_{28}$ are iid $N(\mu, 1 /\tau)$

2.  Assign $(\mu, \tau)$ a vague prior


## Fitting the Model using JAGS

Write a script defining the model.

```{r}
model_string <- "model{
  for (s in 1:28) {
      varhat[s] ~ dgamma(p[s], lam[s])
      p[s] <- n[s] / 2
      lam[s] <- p[s] * tau.betahat[s]
      tau.betahat[s] ~ dgamma(0.001, 0.001)
      sd.betahat[s] <- 1 / sqrt(tau.betahat[s])
      betahat[s] ~ dnorm(beta[s], tau.betahat[s])
      beta[s] ~ dnorm(mu, tau.beta)
  }
  mu ~ dnorm(0.0, 1.0E-6)
  tau.beta ~ dgamma(0.001, 0.001)
  sd.beta <- 1 / sqrt(tau.beta)
  another.beta ~ dnorm(mu, tau.beta)
}"
```

## Define the data.

```{r}
d <- read_csv("data/BBS_survey.csv")
betahat <- d$Trend
varhat <- d$SE ^ 2
n <- d$N_Site
```

## Compile the model.

```{r}
library(rjags)
model <- jags.model(textConnection(model_string), 
                data = list(varhat = varhat,
                            betahat = betahat,
                            n = n))
```

## Running JAGS

Burn-in for 10,000 iterations and run for additional 10,000 iterations, saving the output.

```{r}
update(model, 10000, progress.bar="none")
samp <- coda.samples(model, 
        variable.names=c("beta","mu", "sd.beta"),
          n.iter=10000, progress.bar="none")
```

Create a matrix of simulated draws.

```{r}
S <- data.frame(as.matrix(samp))
```

## Compare the observed and multilevel estimates of trend

```{r, fig.height = 4.5}
Beta <- S[, 1:28]
d1 <- data.frame(Type = "Posterior", N = n,
                 Estimate = apply(Beta, 2, mean))
d2 <- data.frame(Type = "Observed", N = n,
                 Estimate = betahat)
ggplot(rbind(d1, d2), aes(N, Estimate, color=Type)) +
  geom_point()
```

## Posterior of D

```{r, echo = FALSE}
library(TeachBayes)
```

```{r, fig.height = 5.5}
sum.negative <- function(y) {sum(y < 0)}
D <- apply(Beta, 1, sum.negative)
bar_plot(D)
```

## Posterior of rank of R_27 ?

```{r, fig.height = 5.5}
Ranks <- t(apply(Beta, 1, rank))
bar_plot(Ranks[, 27]) + xlim(0, 13)
```


