---
title: "Introduction to JAGS"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

## JAGS

- General program for MCMC simulation of Bayesian hierarchical models

- Similar to WinBUGS/OpenBUGS, but is platform independent (works on Macs)

- Can be run on the command line or run through R by the rjags package

- Write script defining the Bayesian models

- Many distributions are available


## Poisson model

- Assume $y_1, ..., y_n \sim Poisson(\mu)$

- Place a vague prior on $\mu$:

$$
\log \mu \sim N(0, 10000)
$$

- Interested in predicting a future observation $y_{new}$

#### Posterior and Predictive Distributions

- Likelihood:

$$
L(\mu) = \prod_{j=1}^n \frac{\mu^{y_j}\exp(-\mu y_j)}{y_j!}
$$

- Posterior

$$
g(\mu | y) \propto L(\mu) g(\mu)
$$

- Predictive density

$$
f(y^*) = \int g(\mu | y) f(y^* | \mu) \mu
$$

#### JAGS model script

```{r}
modelString = "
model{
for (i in 1:N) {
x[i] ~ dpois(mu)
}
log(mu) <- theta
theta ~ dnorm(0, .0001)
xnew ~ dpois(mu)
}
"
writeLines(modelString, con="poissonmodel.bug")
```

#### Create some fake data

- Assume $\mu = 10$ and $n = 100$

```{r}
set.seed(1234)
N <- 100
x <- rpois(N, 10)
```

#### Set up a model object in JAGS

```{r}
library('rjags')
jags <- jags.model('poissonmodel.bug',
                   data = list('x' = x,
                               'N' = N),
                   n.chains = 1,
                   n.adapt = 100)
```


#### Running JAGS

- Warm up for 1000 iterations

```{r}
update(jags, 1000)
```


- Takes an additional 5000 iterations, storing outcomes ($\mu$ and $y^*$)

```{r}
posterior1 <- coda.samples(jags, c("mu", "xnew"),
                    n.iter=5000, progress.bar="gui")
```


#### Graphs of simulated draws of $\mu$ and $y_{new}$

```{r}
plot(posterior1)
```

#### Summaries

```{r}
summary(posterior1)
```



